Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2020-09-05 spectral radius of matrix new method/paper/conclusion/root.tex
first commit

\documentclass[12pt]{article}
%\usetheme{Warsaw}
\usepackage{times}  % fonts are up to you
\usepackage{amssymb, amsmath, mathrsfs,amsthm}
\usepackage{multicol}
\usepackage{bm}
\usepackage{ulem}

@@

\title{TBA}
\author{Ke-Han Chen\\} % 作者
%\institute[Dep. of A. Math., NCTU]{Department of Applied Mathematics\\National Chiao Tung University}
\date{October 18th, 2018}
\date{\today} % 日期

\maketitle
\begin{abstract}
    This is my abstract.
\end{abstract}

\section{Preliminaries}

\begin{defn}
Spectral radius of a matrix $C$ is the largest absolute value of real eigenvalue of $C$. And here we are interested in spectral radius of the adjacency matrix for some simple graph.
\end{defn}
\begin{defn}
    Eigenvalue of a matrix $C$ corresponds to an eigenvector $u$, where $u$ and its eigenvalue $\lambda $  satisfies $Cu = \lambda u$. 
\end{defn}
    We introduce a notation of sub-matrix, which is taken from some columns and some rows of a matrix. 
    \begin{defn}We use $C[\alpha|\beta]$ to denote the sub-matrix of rows $\alpha$ of $C$ and columns $\beta$ of $C$ and $C(\alpha|\beta)$ is the sub-matrix of rows $[n]$ - $\alpha$ of $C$ and columns $[n] - \beta$ of $C$.
    The bracket [] and parentheses () can be used together is the notation of sub-matrix, such as $C[k|e)$ or $C(k|e]$. 
\end{defn}

\thm
        If $C$ is nonnegative square matrix, then the following holds. \\
        The spectral radius $\rho(C)$ is an eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a corresponding nonnegative left eigenvector.

The following theorem is from [].

\begin{thm}
 Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}
\item[(i)]    $PCQ\leq PC'Q$;
\item[(ii)]  there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $Qu$;
\item[(iii)] there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar $\lambda\in \mathbb{R}$
such that $\lambda$ is an eigenvalue of $C$ with associated  left eigenvector $v^TP$; and
\item[(iv)] $v^TPQu>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
    Moreover, $\lambda=\lambda'$ 
if and only if
    \begin{equation*}
        \label{e3}
(PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
\end{equation*}
\end{thm}

\begin{pf}
            Multiplying the nonnegative vector $u$ in (ii) to the right of both terms of  (i),
        \begin{equation}
            \label{e1}
            PCQu\leq PC'Qu=\lambda'PQu.
        \end{equation}
        Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption (iii) to the left of all terms  in (\ref{e1}), we have
        \begin{equation}
            \label{e2}
            \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
        \end{equation}
        Now delete the positive term $v^TPQu$ by assumption (iv) to obtain $\lambda\leq \lambda'$ and finish the proof of the first part.

        Assume that $\lambda=\lambda'$, so the inequality in (\ref{e2}) is an equality.  Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence, $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with $u_j\not=0.$
         Conversely, (\ref{e3}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=\sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so
            $\lambda=\lambda'$ by (\ref{e2}). \qed

 \end{pf}

 Throughout fix $k\in [n-1]$. Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$ appearing in the  position $k,n$ of $E_{kn}$. We will apply the previous theorem with $P=I$ and $$Q=I+E_{kn}=\begin{pmatrix}
1 &  & & &  & 0 \\
 & 1 &  &      &  &  \\
 &  & \ddots & &  & 1 \\
 &  &        & &  &  \\
  &  & & & 1 &  \\
0 &  & & &  & 1 \\
\end{pmatrix}.$$
\begin{defn}\label{v_rooted}
 ~A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}
\bigskip

The following Lemma is immediate from the above definition.
\bigskip

\begin{lem}
If ~$u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}
\item[(i)] $v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item[(ii)] $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
\qed
\end{lem}

\section{Our first result}

\begin{thm} \label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}
\item[(i)]   $C[-|n)\leq C'[-|n)$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item[(ii)] there exist a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item[(iii)] there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item[(iv)] $v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}
\item[(a)] $c_{ik}+c_{in}=c'_{ik}+c'_{in}$\qquad for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
\item[(b)]
$c'_{ij}=c_{ij}\qquad \hbox{for~}1\leq i\leq n,~1\leq j\leq n-1 \hbox{~with~} v_i\ne 0 \hbox{~and~} v'_j> v'_n.$
\end{enumerate} \qed
\end{thm}

{\bf $k$-rooted matrix}

\begin{defn}[(k,n)-sum]
For a matrix $C'=(c'_{ij})$ of $n$ columns, the $(k, n)$-{\it sum} vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
\end{defn}

\begin{defn}[k-rooted matrix]\label{m_rooted}
A  matrix $C'=(c'_{ij})$ is called {\it $k$-rooted}  if its  columns and its $(k, n)$-sum vector are all $k$-rooted except the last column of $C'$.
\end{defn}

\begin{thm}
Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
    \begin{enumerate}
        \item[(i)]$C'$ is a $k$-rooted matrix, if and only if, $Q^{-1}C'Q$ is nonnegative.
        \item[(ii)]$C'$ has a $k$-rooted eigenvector $v'=Qu$ for $\rho(C')$ as an eigenvalue. $u$ is a nonnegative eigenvector of $Q^{-1}C'Q$ for $\rho(C')$.
        \item[(iii)] $\rho(C')$ = $\rho(Q^{-1}C'Q)$
    \end{enumerate}

\begin{pf}
\subsection{}
(i) is immediate from Definition~\ref{m_rooted} and the observation that 
$$Q^{-1}=I-E_{kn}=\begin{pmatrix}
1 &  & & &  & 0 \\
 & 1 &  &      &  &  \\
 &  & \ddots & &  & -1 \\
 &  &        & &  &  \\
  &  & & & 1 &  \\
0 &  & & &  & 1 \\
\end{pmatrix},$$
and $Q^{-1}C'Q$ is 
$$\begin{pmatrix}
c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
\vdots \\
c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{kn-1}-c'_{nk-1}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
\vdots              & \vdots & \ddots              & \vdots & \vdots \\
c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
\end{pmatrix}.
$$

\subsection{(ii)}

By Lemma~\ref{v_rooted} $v'=Qu$ is $k$-rooted. 
Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have 
$C'Qu=\rho(C')Qu$.

\subsection{(iii)}

Since $C'$ and $Q^{-1}C'Q$ have the same set of eigenvalues, clearly $\rho(C')$ = $\rho(Q^{-1}C'Q)$.

\end{pf}
\end{thm}

\begin{lem}\label{l_diag}
If a square matrix $C'$ has a rooted eigenvector for $\lambda'$, then $C'+dI$ also has
the same rooted eigenvector for $\lambda'+d,$ where $d$ is a constant and $I$ is the identity matrix with the same size of $C'$.
\end{lem}

>>>>

\documentclass{article}

%\usetheme{Warsaw}
\usepackage{times}  % fonts are up to you
\usepackage{amssymb, amsmath, mathrsfs,amsthm}
\usepackage{multicol}
\usepackage{bm}
\usepackage{ulem}

@@

### 2020-09-05 spectral radius of matrix new method/thesis/preliminiaries.tex
first commit

\topmargin=-1.5cm
\oddsidemargin=0.7cm
\textheight=23.5cm
\textwidth=15cm
\setlength{\baselineskip}{24pt}
\renewcommand{\baselinestretch}{1.5} %行

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
%\usepackage{colortbl}

@@

\begin{defn}
    Let $C$ be an $n \times n$ real matrix, and $u \in \mathbb{R}^n$ be a column vector. The scalar $\lambda \in \mathbb{C}$ is an $\it{eigenvalue}$ of C corresponding to an $\it{eigenvector}$ if $Cu = \lambda u$

@@

When $C$ is a real square matrix, the $\textit {spectral radius} $ $\rho(C)$ of $C$ is defined by
$$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

@@

\begin{defn}
Given an undirected graph G, the $\textit{spectral radius}$ of G is the spectral radius of the adjacency matrix A.
\end{defn}

@@

\begin{defn}For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column indices of $C$ respectively,  We use $C[\alpha|\beta]$ to denote the submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$ and $j\in\beta$,
\begin{comment}
$C[\alpha|\beta):=C[\alpha|\overline{\beta}],$ where $\overline{\beta}$ is the complement of $\beta$ in the set of column indices, and
similarly, for the definitions of $C(\alpha|\beta]$ and $C(\alpha|\beta).$
 For $\ell\in \mathbb{N},$ $[\ell]:=\{1, 2, \ldots, \ell\},$ symbol $-$ is the complete set of indices, and we use  $i$ to denote the singleton subset $\{i\}$ to reduce the double use of parentheses. For example of the $n\times n$ matrix $C$,
 $C[-|n)=C[[n]|[n-1]]$ is the $n\times (n-1)$ submatrix of $C$ obtained by deleting the last column of $C$.

@@

\begin{thm}
        If $C$ is a nonnegative square matrix, then spectral radius $\rho(C)$ is an eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a corresponding nonnegative left eigenvector.
\end{thm}

The following theorem is from \cite{chang}.

\begin{thm}
 Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}
\item[(i)]    $PCQ\leq PC'Q$;
\item[(ii)]  there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $Qu$;
\item[(iii)] there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar $\lambda\in \mathbb{R}$
such that $\lambda$ is an eigenvalue of $C$ with associated  left eigenvector $v^TP$; and
\item[(iv)] $v^TPQu>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
    Moreover, $\lambda=\lambda'$
if and only if
    \begin{equation*}
        \label{e3}
(PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
\end{equation*}
\end{thm}

\begin{proof}
            Multiplying the nonnegative vector $u$ in (ii) to the right of both terms of  (i),
        \begin{equation}
            \label{e1}
            PCQu\leq PC'Qu=\lambda'PQu.
        \end{equation}
        Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption (iii) to the left of all terms  in (\ref{e1}), we have
        \begin{equation}
            \label{e2}
            \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
        \end{equation}
        Now delete the positive term $v^TPQu$ by assumption (iv) to obtain $\lambda\leq \lambda'$ and finish the proof of the first part.

        Assume that $\lambda=\lambda'$, so the inequality in (\ref{e2}) is an equality.  Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence, $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with $u_j\not=0.$
         Conversely, (\ref{e3}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=\sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so
            $\lambda=\lambda'$ by (\ref{e2}). \qed

@@

 \newpage
\begin{thebibliography}{20}
\normalsize
\addcontentsline{toc}{chapter}{Bibliography}

\bibitem{chang}
Yen-Jen Chenga, Chih-wen Wenga, A matrix realization of spectral bounds
of the spectral radius of a nonnegative matrix, (journal),....

\bibitem{catalan}
\newblock{James Haglund, The $q$,$t$-Catalan Numbers and the Space of Diagonal Harmonics: With an Appendix on the Combinatorics of Macdonald Polynomials},
\newblock{\it University Lectures Series}, vol. 41, 2008

\bibitem{combinatorics}
\newblock{J.H. van Lint and R.M. Wilson, A Course in Combinatorics, second edition},
\newblock{\it} Cambridge University Press, New York, 2001.

\bibitem{euler}
\newblock{H. W. Gould, Euler's formula nth differences of powers},
\newblock{\it The American Mathematical Monthly }, vol. 85, No. 6, 450-467,1978

\bibitem{rook}
\newblock{Feryal Alayomt and Nicholas Krzywonos, Rook polynomials in three and higher dimensions},
\newblock{\it }, 2009

\bibitem{brualdi}
\newblock{Richard A. Brualdi, Introductory Combinatorics, fifth edition},
\newblock{\it },  Prentice Hall, New Jersey, 2009.

\bibitem{qanalogue}
\newblock{Bruce E. Sagan, Congruence Properties of $q$-Analogs},
\newblock{\it Advances in Mathematics}, 95, 127-143,1992

\bibitem{lagrange}
\newblock{Eric W. Weisstein, Lagrange Interpolating Polynomial}, \newblock{MathWorld}

\bibitem{9}
\newblock{郭子翔,一個新恆等式的發現與研究}, preprint.

\end{thebibliography}

>>>>

\topmargin=-1.5cm
\oddsidemargin=0.7cm
\textheight=23.5cm
\textwidth=15cm
\setlength{\baselineskip}{24pt}
\renewcommand{\baselinestretch}{1.5} %行距

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{comment}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
%\usepackage{colortbl}

@@

\begin{defn}
    Let $C$ be an $n \times n$ real matrix, and $u \in \mathbb{R}^n$ be a nonzero column vector. The scalar $\lambda \in \mathbb{C}$ is an $\it{eigenvalue}$ of C corresponding to the $\it{eigenvector}$ $u,$  if $Cu = \lambda u.$

@@

When $C$ is an $n \times n$ real matrix, the $\textit {spectral radius} $ $\rho(C)$ of $C$ is defined by
$$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

@@

\begin{defn}
Given an undirected graph G, the $\textit{spectral radius}$  $\rho(G) $ of G is the spectral radius of the adjacency matrix of G.
\end{defn}

@@

\begin{defn}For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column indices of $C$ respectively,  We use $C[\alpha|\beta]$ to denote the submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$ and $j\in\beta$,

The following theorem is a part of Perron-Frobenius theorem \cite{'content'}.
\begin{comment}
$C[\alpha|\beta):=C[\alpha|\overline{\beta}],$ where $\overline{\beta}$ is the complement of $\beta$ in the set of column indices, and
similarly, for the definitions of $C(\alpha|\beta]$ and $C(\alpha|\beta).$
 For $\ell\in \mathbb{N},$ $[\ell]:=\{1, 2, \ldots, \ell\},$ symbol $-$ is the complete set of indices, and we use  $i$ to denote the singleton subset $\{i\}$ to reduce the double use of parentheses. For example of the $n\times n$ matrix $C$,
 $C[-|n)=C[[n]|[n-1]]$ is the $n\times (n-1)$ submatrix of $C$ obtained by deleting the last column of $C$.

@@

%Peron Frobius theorem
\begin{thm}
        If $C$ is a nonnegative square matrix, then spectral radius $\rho(C)$ is an eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a corresponding nonnegative left eigenvector.
\end{thm}

The following theorem is from \cite{chang}. For completeness, we also provide the proof.

\begin{thm}
 Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}
\item[(i)]    $PCQ\leq PC'Q$;
\item[(ii)]  there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $Qu$;
\item[(iii)] there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar $\lambda\in \mathbb{R}$
such that $\lambda$ is an eigenvalue of $C$ with associated  left eigenvector $v^TP$; and
\item[(iv)] $v^TPQu>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
    Moreover, $\lambda=\lambda'$
if and only if
    \begin{equation}
        \label{pre0}
(PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
\end{equation}
\end{thm}

\begin{proof}
            Multiplying the nonnegative vector $u$ in assumption (ii), where $Qu$ is eigenvector of $C'$,  to the right of both terms of  (i),
        \begin{equation}
            \label{pre1}
            PCQu\leq PC'Qu=\lambda'PQu.
        \end{equation}
        Multiplying the nonnegative vector $v^T$  in assumption (iii) , where $v^TP$ is left eigenvector of $C$ for $\lambda$,to the left of all terms  in (\ref{pre1}), we have
        \begin{equation}
            \label{pre2}
            \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
        \end{equation}
        Now delete the positive term $v^TPQu$ by assumption (iv) to obtain $\lambda\leq \lambda'$ and finish the proof of the first part.

        Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.  Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence, $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with $u_j\not=0.$
         Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=\sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so
            $\lambda=\lambda'$ by (\ref{pre2}).

@@
